The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2021-09-26 paper/paper-jsdse/paper-jsdse.tex
comparison plots
comparison

In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data. Referring to Loo and Jonge (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data. The raw demographic and employment data from the database are not in a tidy format. Hence, we perform data wrangling to get the input data. We inspect the input data to find eligible observations to the target data. One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey. This needs us to join the extracted demographic and employment data.

Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value.
From this stage forward, we assume that we already have valid data. We also subset the data to get the wages of high school dropouts. We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called \texttt{yowie}. A visualisation of these process is displayed in Figure \ref{fig:flow-chart}.

@@

@@

Figure \ref{fig:plotting-sw-do} shows the longitudinal profiles of the two sets. (A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data. This suggests some individuals were removed in the original data. Overall, the profiles show some similarity.

>>>>

Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (Loo and Jonge 2018) to create a refreshed wages data set.

@@

\begin{verbatim}
#> # A tibble: 3 x 2
#>   race         n
#>   <fct>    <int>
#> 1 black      246
#> 2 hispanic   204
#> 3 white      438
\end{verbatim}

\begin{verbatim}
#> # A tibble: 2 x 2
#>   hgc       n
#>   <chr> <int>
#> 1 12TH    522
#> 2 8TH     366
\end{verbatim}

\begin{verbatim}
#> # A tibble: 3 x 2
#>   race                        n
#>   <fct>                   <int>
#> 1 HISPANIC                  219
#> 2 BLACK                     346
#> 3 NON-BLACK, NON-HISPANIC   623
\end{verbatim}

\begin{verbatim}
#> # A tibble: 2 x 2
#>   hgc12          n
#>   <chr>      <int>
#> 1 12TH         967
#> 2 BELOW 12TH   221
\end{verbatim}

@@

Figure \ref{fig:plotting-sw-do} shows a comparison of the two sets. (A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data. This suggests some individuals were removed in the original data.
In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in tne refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced. On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original. In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.

### 2021-09-26 paper/paper.Rmd
comparison plots
comparison

```{r compare-plot, fig.cap="Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals. A robust linear model prediction was used to correct mean hourly wages value. We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed. Some spikes might remain when wage vaues are plausible.", fig.height=8, fig.width=10, out.width="100%"}
```

@@

In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data. Referring to @LooMarkvander2018Sdcw, this section shows  part of statistical value chain, which is data cleaning from raw, input, and valid data. The raw demographic and employment data from the database are not in a tidy format. Hence, we perform data wrangling to get the input data. We inspect the input data to find eligible observations to the target data. One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey. This needs us to join the extracted demographic and employment data.

Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value. 
From this stage forward, we assume that we already have valid data. We also subset the data to get the wages of high school dropouts. We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called `yowie`. A visualisation of these process is displayed in  Figure \ref{fig:flow-chart}.

>>>>

```{r compare-plot, fig.cap="Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals. A robust linear model prediction was used to correct mean hourly wages value. We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed. Some spikes might remain when wage vaues are plausible.", fig.height=8, fig.width=10, out.width="90%"}
```

@@

Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data [@LooMarkvander2018Sdcw] to create a refreshed wages data set.

<!--To summarise, the  a workflow from getting the data from the NSLY79 database until a ready-analysed data. Referring to @LooMarkvander2018Sdcw, this section shows  part of statistical value chain, which is data cleaning from raw, input, and valid data. The raw demographic and employment data from the database are not in a tidy format. Hence, we perform data wrangling to get the input data. We inspect the input data to find eligible observations to the target data. One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey. This needs us to join the extracted demographic and employment data.

Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value. 
From this stage forward, we assume that we already have valid data. We also subset the data to get the wages of high school dropouts. We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called `yowie`. A visualisation of these process is displayed in  Figure \ref{fig:flow-chart}.-->
